Quantifying Roughness in SAR Imagery with the Rényi Entropy

### 2025-03-29 Text/Articles/RENYI_ENTROPY_SAR_2025/R0-Renyi-entropy-sar.qmd
Submitted version

      Data and code are available at: <https://github.com/rjaneth/renyi-entropy> 
    # bio: |
    #   Use `IEEEbiographynophoto` and the author name
    #   as the argument followed by the biography text.
    # note: "Template created June 23, 2023; revised `r format(Sys.Date(),format='%B %d, %Y')`."
abstract: |
  Quantifying surface roughness in synthetic aperture radar (SAR) data is critical for accurate geophysical interpretation and remote sensing applications. 
  We propose a test statistic based on a non-parametric estimation of Rényi entropy to characterize surface roughness from SAR intensity data. 
  The statistic is refined using bootstrap to improve its stability, size and power. 
  This approach enhances roughness quantification by capturing scale-dependent variations and addressing data-driven uncertainty.
  Experimental results demonstrate the robustness of the proposed method in distinguishing roughness patterns, offering a statistically rigorous tool for SAR-based terrain analysis.
  <!-- Synthetic aperture radar (SAR) systems have already been successfully used to solve remote sensing problems.  -->
  <!-- A disadvantage of SAR images is the presence of speckle, which is fully developed in homogeneous areas and gamma-distributed in these scenes.  -->
  <!-- In heterogeneous areas, the intensity values are $\mathcal{G}^0_I$-distributed.  -->
  <!-- In this way, the identification of roughness SAR areas (as opposed to homogeneous ones) is an important task.  -->
  <!-- In this work, we propose a family of hypothesis tests driven by an order parameter to identify roughness features in SAR intensity data using Rényi entropy.  -->
  <!-- In particular, we use a non-parametric estimator for the Rényi entropy and investigate some of its properties.  -->
  <!-- As a practical evaluation method, we develop $p$-value maps on which one can observe both (i) the heterogeneous evidence change per texture and (ii) the prediction of homogeneous and heterogeneous categories.  -->
  <!-- The results are in favor of the Rényi-based heterogeneity detector compared to the one based on Shannon entropy. -->
keywords: [Rényi entropy, Gamma distribution, heterogeneity, SAR, hypothesis tests, bootstrap]

@@

```
\renewcommand{\tablename}{TABLE}
# Introduction
[S]{.IEEEPARstart}[ynthetic]{}
 Aperture Radar (SAR) technology has become essential for a wide range of applications&nbsp;[@Yu2023;@Akbarizadeh2012;@Mondini2021]. 
 It provides high-resolution data that is independent of sunlight and operates in a variety of weather conditions that facilitate global Earth monitoring&nbsp;[@Zeng2020].
Despite its advantages, the effective use of SAR data depends on understanding its statistical properties because this data is affected by speckle, a noise-like interference effect&nbsp;[@Argenti2013; @Choi2019; @Baraha2023]. In intensity format, speckle is typically non-Gaussian, and its presence complicates subsequent image analysis tasks.

The $\mathcal{G}^0_I$ distribution effectively characterizes SAR intensity data because it captures different levels of roughness. 
A limiting case is the Gamma distribution, which arises when speckle is fully developed, indicating textureless regions. 
Although these assumptions offer flexibility in describing SAR intensities, selecting the appropriate distribution can be challenging for two main reasons: the small sample sizes used in practical applications and the inherent difficulties associated with parameter estimation. 
These challenges complicate model selection and highlight the need to explore alternative statistical approaches.

@@

We propose a statistical test based on a non-parametric estimator of Rényi entropy to identify heterogeneous regions in SAR data.
The test assesses whether the observed Rényi entropy significantly differs from its expected theoretical value under the assumption of homogeneity.
Compared to our previous approach using Shannon entropy&nbsp;[@Frery2024], the Rényi-based approach improves the detection of heterogeneity.

@@

The main distributions considered for SAR intensity data are the $\Gamma_{\text{SAR}}$ distribution, which is suitable for fully developed speckle, and the $\mathcal{G}^0_I$ distribution, which is able to describe roughness&nbsp;[@Frery1997]. These distributions are characterized by the following probability density functions (pdfs):
\begin{equation}
	f_{\Gamma_{\text{SAR}}}\bigl(z;L, \mu \bigr) 
    = \frac{L^L}{\Gamma(L)\,\mu^L} z^{L-1} 
    \exp \biggl(-\frac{Lz}{\mu}\biggr)
    \mathbbm 1_{\mathbbm R_+}(z), \label{E:gamma1}
\end{equation}
and
\begin{multline}
    f_{\mathcal{G}^0_I}\bigl(z; \mu, \alpha, L \bigr) 
    = \frac{L^L\,\Gamma(L-\alpha)}
    {\bigl[-\mu(\alpha+1)\bigr]^{\alpha} \Gamma(-\alpha)\,\Gamma(L)}\\
    \frac{z^{L-1}}
    {\bigl[-\mu(\alpha+1)+Lz\bigr]^{L-\alpha}}
    \mathbbm 1_{\mathbbm R_+}(z), \label{E:gi01}
\end{multline}
where $\mu > 0$ is the mean,
$\alpha < 0$ measures the roughness, $L \geq 1$ is the number of
looks, $\Gamma(\cdot)$ is the gamma function, and
$\mathbbm 1_{A}(z)$ is the indicator function of the set $A$. 
As demonstrated by&nbsp;[@Frery1997], the $\Gamma_{\text{SAR}}$  model is a particular case of the $\mathcal{G}^0_I$ distribution. Specifically, for a given $\mu$ fixed,
$$
f_{\mathcal{G}^0_I}\big(z; \mu, \alpha, L\big)
\longrightarrow 
f_{\Gamma_{\text{SAR}}}(z;L, \mu) 
$$
when $\alpha\to-\infty$.

@@

Introduced by Alfréd Rényi in 1961&nbsp;[@renyi1961measures], this measure generalizes several well-known entropies, including Shannon's&nbsp;[@Ribeiro2021]. 
For a continuous random variable $Z$ with pdf $f(z)$, the Rényi entropy of order $\lambda \in \mathbbm R_+ \setminus \{1\}$, is defined as:
\begin{equation}
\label{E:entropy2}
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.
\end{equation}
Using&nbsp;\eqref{E:entropy2}, we derive closed-form expressions for the Rényi entropy of the $\Gamma_{\mathrm{SAR}}$ and the $\mathcal{G}^0_I$ distributions:
\begin{multline}
\label{eq-HGammaSAR}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr)
= 
\ln \mu - \ln L + \frac{1}{1-\lambda}
\Bigl[
  -\lambda\,\ln\Gamma(L)  \\ + \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  - \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\Bigr],
\end{multline}
and
\begin{multline}  
\label{eq-HGI0}  
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr) + \ln(-1 - \alpha) \\  
+ \frac{1}{1 - \lambda} \Bigl[ \lambda\bigl(\ln\Gamma(L - \alpha) - \ln\Gamma(-\alpha)\bigr) \\ 
+ \ln\Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr) - \ln\Gamma\bigl(\lambda(L - \alpha)\bigr) \\
+ \bigl(\lambda(L-1)+1\bigr)\ln\lambda \Bigr].  
\end{multline}
The derivation of the latter result is proven in the Appendix.
Note that the Rényi entropy of $\mathcal{G}^0_I$ can be expressed in terms of that of $\Gamma_{\mathrm{SAR}}$, with additional terms depending on the roughness parameter $\alpha$. As $\alpha \to -\infty$, these terms tend to zero, reducing&nbsp;\eqref{eq-HGI0} to&nbsp;\eqref{eq-HGammaSAR}.

@@

In order to estimate&nbsp;\eqref{E:entropy2} from a random sample, one may estimate the parameters that index the probability density function $f$, and integrate.
This approach heavily relies on the model assumptions and on the absence of outliers.
Alternatively, one may estimate directly $f$.
Such non-parametric estimation of $H(Z)$ has been widely studied using spacing-based estimators, which rely on differences between order statistics&nbsp;[@vasicek1976test;@Bert1992; @Ebrahimi1994; @Wieczorkowski1999; @IbrahimAlOmari2014]. 
Recently, Al-Labadi et al.&nbsp;[@AlLabadi2024] proposed a non-parametric estimator for Rényi entropy following this approach.

@@

We aim to determine the optimal order $\lambda$ for the Rényi entropy estimator for a sample size $n=49$.
To identify this optimal value, we analyze both the mean squared error (MSE) and bias of the estimator across different values of $\lambda$. Lower MSE and bias indicate better performance of the estimator in approximating the true entropy.

We found that for $L > 1$ the optimal value is $\lambda = 0.9$, as it minimizes the MSE while maintaining a low bias.
However, for $L = 1$, the optimal $\lambda$ tends to be higher and we chose $\lambda = 3$ to achieve good results. 
@fig-plotf illustrates the case for $L = 5$.
```{r fig-plotf, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="42%",  fig.pos="hbt", fig.cap="Bias and MSE as a function of $\\lambda$, with $n=49$, $L=5$.", fig.width=7, fig.height=4.5}

@@

```
For subsequent simulations we use the $\widetilde{H}_{\lambda}$ estimator.

@@

Since homogeneity corresponds to $\alpha \to -\infty$ in the $\mathcal{G}^0_I$ model, classical inference methods are inapplicable. 
Instead, we propose a test statistic based on the Rényi entropy estimator, comparing it with its theoretical expectation under $\mathcal{H}_0$ to distinguish between homogeneous and heterogeneous areas.

@@

Assume we have an estimator $\widetilde{H}_{\lambda}$ for the Rényi entropy of an arbitrary model. For testing&nbsp;\eqref{eq:hypothesis_test}, this estimator is expected to be close to the theoretical expression given in Equation&nbsp;\eqref{eq-HGammaSAR}. 
Since $L\geq1$ is known, we define the test statistic as follows:
\begin{multline}
\label{eq-test}
S_{\widetilde{H}_{\lambda}}(\bm{Z}; L) = \widetilde{H}_{\lambda} - \bigl\{\ln \widehat{\mu} - \ln L + \frac{1}{1-\lambda}
\bigl[-\lambda\,\ln\Gamma(L) \\ 
+ \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  
- \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\bigr]\bigr\},
\end{multline}
where $\widehat{\mu}={n}^{-1}\sum_{i=1}^n Z_{i}$ is the sample mean.
When the null hypothesis holds, this test statistic is expected to be close to zero.
@fig-density_entropyR shows the empirical density of $S_{\widetilde{H}_{\lambda}}(\bm{Z}; L)$, based on $10^4$ replications of the $\Gamma_{\text{SAR}}$ model for each sample size $n \in \{49,81, 121\}$ with $\lambda=0.9$ and $L \in \{5,18\}$.
<!-- ACF 10^4 replications of samples of size?   -->
<!-- Corrected -->
```{r Simulated_densityR, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

@@

We define the standardized test statistic:
\begin{equation*}
\varepsilon = \frac{S_{\widetilde{H}_{\lambda}}(\bm{Z}; L) - \hat{\mu}_S}{\hat{\sigma}_S},
\end{equation*}
where $\hat{\mu}_S$ and $\hat{\sigma}_S$ are the empirical mean and standard deviation of $S_{\widetilde{H}_{\lambda}}(\bm{Z}; L)$, obtained by Monte Carlo simulations under the null hypothesis. 
By the central limit theorem, $\varepsilon$ asymptotically follows a standard normal distribution:
\begin{equation*}
\varepsilon \overset{d}{\longrightarrow} \mathcal{N}(0,1)\,.
\end{equation*}
Thus, the $p$-values are calculated as $2\Phi(-|\varepsilon|)$,
where $\Phi(\cdot)$ is the cumulative distribution function of the standard normal distribution.
@fig-density_entropyR_standardized0 shows smoothed histograms of standardized test statistics.

@@

Further, we examined the power under the alternative hypothesis ($\mathcal{G}^0_I$ distribution) with $\mu=1$, $\lambda=0.9$ and $\alpha=-2$.
As expected, power improves as the sample size and number of looks increase, demonstrating the test effectiveness. The results are shown in Table&nbsp;\ref{tab:table_size_power}.
```{r Simulated_error_type_I, echo=FALSE, message=FALSE}
if (!file.exists("./Data/type_I_results.Rdata")) {

@@

The $p$-value maps use a color gradient, where darker regions indicate areas with higher roughness and lighter regions denote smoother, less textured surfaces.
The binary maps classify these results: white areas represent $p$-values greater than $0.05$, indicating no statistical evidence to reject the null hypothesis (homogeneous regions). 
In contrast, black areas represent regions with $p$-values below $0.05$, providing statistical evidence to reject the null hypothesis and indicating heterogeneity.

The Rényi entropy-based test demonstrated greater sensitivity in detecting textural variations compared to the Shannon-based test, due to the flexibility provided by the parameter $\lambda$, as observed in the binary maps in Figs.&nbsp;\ref{fig:london-c},&nbsp;\ref{fig:munich-c}, and&nbsp;\ref{fig:dublin-c}.
<!-- ACF Where is the Shannon-based approach shown? -->
<!-- JA: The results of the Shannon-based approach can be observed in the corresponding binary maps. -->
```{=latex}
\begin{figure*}[hbt]
    \centering
    \begin{subfigure}{0.17\textwidth}
        \includegraphics[width=\linewidth]{./Figures/london_2000.png}
        \caption{ SAR image}
        \label{fig:london-a}
    \end{subfigure}
    \begin{subfigure}{0.22\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_pvalue_london_renyi.png}
        \caption{$p$-value map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:london-b}
    \end{subfigure}
    \begin{subfigure}{0.17\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_005_london_renyi_L1_.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:london-c}
    \end{subfigure}
    \begin{subfigure}{0.22\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_pvalue_london_Shannon_c1.png}
        \caption{$p$-value map for $\tiny{S_{\widetilde{H}_{\text{AO}}}}$ }
        \label{fig:london-d}
    \end{subfigure}
    \begin{subfigure}{0.17\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_005_london_shannon.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:london-e}
    \end{subfigure}
    \caption{Detection of heterogeneous areas in a SAR image over London, UK: comparison with the tests $\small{S_{\widetilde{H}_{\lambda}}}$ and $\small{S_{\widetilde{H}_{\text{AO}}}}$.  }
    \label{fig:london}
\end{figure*}

@@

This study presented a new statistical approach based on Rényi entropy for detecting heterogeneity in SAR images, distinguishing between fully developed speckle and heterogeneous clutter. 
The statistics of the proposed test was computed non-parametrically and its performance was evaluated by a Monte Carlo study. 
The results showed that the method effectively controls the probability of Type&nbsp;I error while maintaining a high detection performance that improves as the sample size increases.

<!-- ACF Again, where are the Shannon results? -->
<!-- JA: The Shannon results are visually represented through p-value and binary maps for each image -->
The effectiveness of the Rényi entropy-based test in SAR images was compared with the Shannon entropy-based test using $p$-value and binary maps. 
The differences between the two methods are evident in the binary maps at a $0.05$ significance level.
Although both tests successfully detected heterogeneous regions, the Rényi-based test performed better due to its flexibility in adjusting the $\lambda$ parameter, which allowed for finer differentiation of texture variations.

>>>>

      Data and code are available at: <https://github.com/rjaneth/renyi-entropy> 
    # bio: |
    #   Use `IEEEbiographynophoto` and the author name
    #   as the argument followed by the biography text.
    # note: "Template created June 23, 2023; revised `r format(Sys.Date(),format='%B %d, %Y')`."
abstract: |
  Quantifying surface roughness in synthetic aperture radar (SAR) data is critical for accurate geophysical interpretation and remote sensing applications. 
  We propose a test statistic based on a non-parametric estimation of Rényi entropy to characterize surface roughness from SAR intensity data. 
  The statistic is refined using bootstrap to improve its stability, size, and power. 
  This approach enhances roughness quantification by capturing scale-dependent variations and addressing data-driven uncertainty.
  Experimental results demonstrate the robustness of the proposed method in distinguishing roughness patterns, offering a statistically rigorous tool for SAR-based terrain analysis.
  <!-- Synthetic aperture radar (SAR) systems have already been successfully used to solve remote sensing problems.  -->
  <!-- A disadvantage of SAR images is the presence of speckle, which is fully developed in homogeneous areas and gamma-distributed in these scenes.  -->
  <!-- In heterogeneous areas, the intensity values are $\mathcal{G}^0_I$-distributed.  -->
  <!-- In this way, the identification of roughness SAR areas (as opposed to homogeneous ones) is an important task.  -->
  <!-- In this work, we propose a family of hypothesis tests driven by an order parameter to identify roughness features in SAR intensity data using Rényi entropy.  -->
  <!-- In particular, we use a non-parametric estimator for the Rényi entropy and investigate some of its properties.  -->
  <!-- As a practical evaluation method, we develop $p$-value maps on which one can observe both (i) the heterogeneous evidence change per texture and (ii) the prediction of homogeneous and heterogeneous categories.  -->
  <!-- The results are in favor of the Rényi-based heterogeneity detector compared to the one based on Shannon entropy. -->
keywords: [Rényi entropy, Gamma distribution, heterogeneity, SAR, hypothesis tests, bootstrap]

@@

```
\renewcommand{\tablename}{TABLE}
# Introduction
[S]{.IEEEPARstart}[ynthetic]{}
 Aperture Radar (SAR) technology has become essential for many applications&nbsp;[@Yu2023;@Akbarizadeh2012;@Mondini2021]. 
 It provides high-resolution data independent of sunlight and operates in various weather conditions that facilitate global Earth monitoring&nbsp;[@Zeng2020].
Despite its advantages, the effective use of SAR data depends on understanding its statistical properties because it is affected by speckle, a noise-like interference effect&nbsp;[@Argenti2013; @Choi2019; @Baraha2023]. In intensity format, speckle is typically non-Gaussian, and its presence complicates subsequent image analysis tasks.

The $\mathcal{G}^0_I$ distribution effectively characterizes SAR intensity data because it captures different roughness levels. 
A limiting case is the Gamma distribution, which arises when speckle is fully developed, indicating textureless regions. 
Although these assumptions offer flexibility in describing SAR intensities, selecting the appropriate distribution can be challenging for two reasons: the small sample sizes used in practical applications and the inherent difficulties associated with parameter estimation. 
These challenges complicate model selection and highlight the need to explore alternative statistical approaches.

@@

We propose a statistical test based on a non-parametric estimator of Rényi entropy to identify heterogeneous regions in SAR data.
The test assesses whether the observed Rényi entropy significantly differs from its expected theoretical value under the homogeneity assumption.
Compared to our previous approach using Shannon entropy&nbsp;[@Frery2024], the Rényi-based approach improves the detection of heterogeneity.

@@

The main distributions considered for SAR intensity data are the $\Gamma_{\text{SAR}}$ distribution, which is suitable for fully developed speckle, and the $\mathcal{G}^0_I$ distribution, which can describe roughness&nbsp;[@Frery1997]. These distributions are characterized by the following probability density functions (pdfs):
\begin{equation}
	f_{\Gamma_{\text{SAR}}}\bigl(z;L, \mu \bigr) 
    = \frac{L^L}{\Gamma(L)\,\mu^L} z^{L-1} 
    \exp \biggl(-\frac{Lz}{\mu}\biggr)
    \mathbbm 1_{\mathbbm R_+}(z), \label{E:gamma1}
\end{equation}
and
\begin{multline}
    f_{\mathcal{G}^0_I}\bigl(z; \mu, \alpha, L \bigr) 
    = \frac{L^L\,\Gamma(L-\alpha)}
    {\bigl[-\mu(\alpha+1)\bigr]^{\alpha} \Gamma(-\alpha)\,\Gamma(L)}\\
    \frac{z^{L-1}}
    {\bigl[-\mu(\alpha+1)+Lz\bigr]^{L-\alpha}}
    \mathbbm 1_{\mathbbm R_+}(z), \label{E:gi01}
\end{multline}
where $\mu > 0$ is the mean,
$\alpha < 0$ measures the roughness, $L \geq 1$ is the number of
looks, $\Gamma(\cdot)$ is the gamma function, and
$\mathbbm 1_{A}(z)$ is the indicator function of the set $A$. 
The $\Gamma_{\text{SAR}}$  model is a particular case of the $\mathcal{G}^0_I$ distribution&nbsp;[@Frery1997]: for $\mu$ fixed,
$$
f_{\mathcal{G}^0_I}\big(z; \mu, \alpha, L\big)
\longrightarrow 
f_{\Gamma_{\text{SAR}}}(z;L, \mu) \text{ when } \alpha\to-\infty.
$$

@@

Introduced by Alfréd Rényi in 1961&nbsp;[@renyi1961measures], this measure generalizes several well-known entropies, including Shannon's&nbsp;[@Ribeiro2021]. 
For a continuous random variable $Z$ with pdf $f(z)$, the Rényi entropy of order $\lambda \in \mathbbm R_+ \setminus \{1\}$, is defined as:
\begin{equation}
\label{E:entropy2}
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.
\end{equation}
Using&nbsp;\eqref{E:entropy2}, we derive closed-form expressions for the Rényi entropy of the $\Gamma_{\mathrm{SAR}}$ and the $\mathcal{G}^0_I$ distributions:
\begin{multline}
\label{eq-HGammaSAR}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr)
= 
\ln \mu - \ln L + \frac{1}{1-\lambda}
\Bigl[
  -\lambda\,\ln\Gamma(L)  \\ + \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  - \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\Bigr],
\end{multline}
and
\begin{multline}  
\label{eq-HGI0}  
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr) + \\ \ln(-1 - \alpha)  
+ \frac{1}{1 - \lambda} \Bigl[ \lambda\bigl(\ln\Gamma(L - \alpha) - \ln\Gamma(-\alpha)\bigr) \\ 
+ \ln\Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr) - \ln\Gamma\bigl(\lambda(L - \alpha)\bigr) \\
+ \bigl(\lambda(L-1)+1\bigr)\ln\lambda \Bigr].  
\end{multline}
The derivation of the latter result is proven in the Appendix.
Note that the Rényi entropy of $\mathcal{G}^0_I$ can be expressed in terms of that of $\Gamma_{\mathrm{SAR}}$, with additional terms that depend on the roughness parameter $\alpha$. As $\alpha \to -\infty$, these terms tend to zero, reducing&nbsp;\eqref{eq-HGI0} to&nbsp;\eqref{eq-HGammaSAR}.
This work's core idea is to estimate this difference and check if it is significantly different from zero.

@@

In order to estimate&nbsp;\eqref{E:entropy2} from a random sample, one may estimate the parameters that index the probability density function $f$ and integrate.
This approach relies heavily on the model assumptions and the absence of outliers.
Alternatively, one may estimate directly $f$.
Such non-parametric estimation of $H(Z)$ has been widely studied using spacing-based estimators, which rely on differences between order statistics&nbsp;[@vasicek1976test;@Bert1992; @Ebrahimi1994; @Wieczorkowski1999; @IbrahimAlOmari2014]. 
Recently, Al-Labadi et al.&nbsp;[@AlLabadi2024] proposed a non-parametric estimator for Rényi entropy following this approach.

@@

We aim to determine the optimal order $\lambda$ for the Rényi entropy estimator for a sample size $n=49$.
To identify this optimal value, we analyze both the mean squared error (MSE) and bias of the estimator across different values of $\lambda$. Lower MSE and bias indicate better estimator performance in approximating the true entropy.

We found that the optimal value for $L > 1$ is $\lambda = 0.9$, as it minimizes the MSE while maintaining a low bias.
However, for $L = 1$, the optimal $\lambda$ tends to be higher, and we chose $\lambda = 3$ to achieve good results. 
@fig-plotf illustrates the case for $L = 5$.
```{r fig-plotf, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="42%",  fig.pos="hbt", fig.cap="Bias and MSE as a function of $\\lambda$, with $n=49$, $L=5$.", fig.width=7, fig.height=4.5}

@@

```
We use the $\widetilde{H}_{\lambda}$ estimator for subsequent simulations.

@@

Classical inference methods are inapplicable since homogeneity corresponds to $\alpha \to -\infty$ in the $\mathcal{G}^0_I$ model. 
Instead, we propose a test statistic based on the Rényi entropy estimator, comparing it with its theoretical expectation under $\mathcal{H}_0$ to distinguish between homogeneous and heterogeneous areas.

@@

Assume we have an estimator $\widetilde{H}_{\lambda}$ for the Rényi entropy of an arbitrary model. For testing&nbsp;\eqref{eq:hypothesis_test}, this estimator is expected to be close to the theoretical expression given in Equation&nbsp;\eqref{eq-HGammaSAR}. 
Since $L\geq1$ is known, we define the test statistic as follows:
\begin{multline}
\label{eq-test}
S_{\widetilde{H}_{\lambda}}(\bm{Z}; L) = \widetilde{H}_{\lambda} - \bigl\{\ln \widehat{\mu} - \ln L + \frac{1}{1-\lambda}
\bigl[-\lambda\,\ln\Gamma(L) \\ 
+ \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  
- \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\bigr]\bigr\},
\end{multline}
where $\widehat{\mu}={n}^{-1}\sum_{i=1}^n Z_{i}$ is the sample mean.
This test statistic should be close to zero when the null hypothesis holds.
@fig-density_entropyR shows the empirical density of $S_{\widetilde{H}_{\lambda}}(\bm{Z}; L)$, based on $10^4$ replications of the $\Gamma_{\text{SAR}}$ model for each sample size $n \in \{49,81, 121\}$ with $\lambda=0.9$ and $L \in \{5,18\}$.
<!-- ACF 10^4 replications of samples of size?   -->
<!-- Corrected -->
```{r Simulated_densityR, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

@@

The standardized test statistic 
$\varepsilon = ({S_{\widetilde{H}_{\lambda}}(\bm{Z}; L) - \hat{\mu}_S})/{\hat{\sigma}_S}$ is asymptotically standard normal distributed,
where $\hat{\mu}_S$ and $\hat{\sigma}_S$ are the empirical mean and standard deviation of $S_{\widetilde{H}_{\lambda}}(\bm{Z}; L)$, obtained by Monte Carlo simulations under the null hypothesis. 
Thus, the $p$-values are calculated as $2\Phi(-|\varepsilon|)$,
where $\Phi(\cdot)$ is the cumulative distribution function of the standard normal distribution.
@fig-density_entropyR_standardized0 shows smoothed histograms of standardized test statistics.

@@

Further, we examined the power under the alternative hypothesis ($\mathcal{G}^0_I$ distribution) with $\mu=1$, $\lambda=0.9$ and $\alpha=-2$.
As expected, power improves as the sample size and the number of looks increase, demonstrating the test effectiveness. The results are shown in Table&nbsp;\ref{tab:table_size_power}.
```{r Simulated_error_type_I, echo=FALSE, message=FALSE}
if (!file.exists("./Data/type_I_results.Rdata")) {

@@

The $p$-value maps use a color gradient, where darker regions indicate areas with higher roughness, and lighter regions denote smoother, less textured surfaces.
The binary maps classify these results: white areas represent $p$-values greater than $0.05$, indicating no statistical evidence to reject the null hypothesis (homogeneous regions). 
In contrast, black areas represent regions with $p$-values below $0.05$, providing statistical evidence to reject the null hypothesis and indicating heterogeneity.

The Rényi entropy-based test demonstrated greater sensitivity in detecting textural variations than the Shannon-based test due to the flexibility provided by the parameter $\lambda$, as observed in the binary maps in Figs.&nbsp;\ref{fig:london-c},&nbsp;\ref{fig:munich-c}, and&nbsp;\ref{fig:dublin-c}.
<!-- ACF Where is the Shannon-based approach shown? -->
<!-- JA: The results of the Shannon-based approach can be observed in the corresponding binary maps. -->
```{=latex}
\begin{figure*}[hbt]
    \centering
    \begin{subfigure}{0.17\textwidth}
        \includegraphics[width=\linewidth]{./Figures/london_2000.png}
        \caption{ SAR image}
        \label{fig:london-a}
    \end{subfigure}
    \begin{subfigure}{0.22\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_pvalue_london_renyi.png}
        \caption{$p$-value map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:london-b}
    \end{subfigure}
    \begin{subfigure}{0.17\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_005_london_renyi_L1_.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:london-c}
    \end{subfigure}
    \begin{subfigure}{0.22\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_pvalue_london_Shannon_c1.png}
        \caption{$p$-value map for $\tiny{S_{\widetilde{H}_{\text{AO}}}}$ }
        \label{fig:london-d}
    \end{subfigure}
    \begin{subfigure}{0.17\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_005_london_shannon.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:london-e}
    \end{subfigure}
    \caption{Detection of heterogeneous areas in a SAR image over London, UK: comparison with the tests $\small{S_{\widetilde{H}_{\lambda}}}$ and $\small{S_{\widetilde{H}_{\text{AO}}}}$.  }
    \label{fig:london}
\end{figure*}

@@

This study presented a new statistical approach based on Rényi entropy for detecting heterogeneity in SAR images, distinguishing between fully developed speckle and heterogeneous clutter. 
The proposed test performance was evaluated by a Monte Carlo study. 
The results showed that the method effectively controls the probability of Type&nbsp;I error while maintaining a high detection performance that improves as the sample size increases.

<!-- ACF Again, where are the Shannon results? -->
<!-- JA: The Shannon results are visually represented through p-value and binary maps for each image -->
The effectiveness of the Rényi entropy-based test in SAR images was compared with the Shannon entropy-based test using $p$-value and binary maps. 
The differences between the two methods are evident in the binary maps at a $0.05$ significance level.
Although both tests successfully detected heterogeneous regions, the Rényi-based test performed better due to its flexibility in adjusting the $\lambda$ parameter, allowing for finer texture variation differentiation.
